---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tjm.praat

<!-- badges: start -->
<!-- badges: end -->

The goal of tjm.praat is to make it easier to run
[Praat](http://www.fon.hum.uva.nl/praat/) scripts as part of an R workflow. This
package provides one main function `wrap_praat_script()`, plus some scripts
designed to work with this function. The name `tjm.praat` indicates that the
package houses my (tjm's) `.praat` files.

## Installation

You can install tjm.praat from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tjmahr/tjm.praat")
```

## Example of wrapping a Praat script as a function

Let's make a function that draws a Praat textgrid and saves the image as `.png`
file. First, let's make the minimal working Praat script, and store it as a
string in R. We use `glue::glue()` because it cleans up the leading and trailing
blank lines and the indentations on the string.

```{r}
script <- glue::glue(
  '
  form Draw a textgrid
    sentence Textgrid_in
    integer Width 6
    integer Height 4
    sentence Png_out
  endform

  Read from file: textgrid_in$
  Select outer viewport: 0, width, 0, height
  
  Draw: 0, 0, "yes", "yes", "yes"
  Save as 300-dpi PNG file: png_out$
  '
)
```

Before we can wrap the script into a function, we need to set the **Praat 
location**. By default, tjm.praat searches for a location in 
`options("tjm.praat_location")` and `Sys.which("praat")`. So, on my machine, 
tjm.praat manages to find a Praat executable:

```{r}
library(tjm.praat)
get_praat_location()
```

But if I want to set or override this location, I can set the location for the 
R session. For example, here I point to the copy of Praat.exe on my desktop.

```{r}
# Trying to use a relative path for the demo. 
# (These are normalized anyways).
set_praat_location("~/../Desktop/Praat.exe")

get_praat_location()
```


Now, we can convert this script into an R function. We tell
`wrap_praat_script()` to return the last argument of the script (`png_out$` in
this example) back to R after the script runs.

```{r}
f_draw_textgrid <- wrap_praat_script(
  script_code_to_run = script,
  return = "last-argument"
)
```

Now we set up the file arguments to script. First, we need a textgrid file.
Let's use a textgrid bundled with the package. The textgrid shows the results of
a forced-alignment on the phrase "bird house".

```{r}
tg_in <- system.file(
  "demo-textgrids/birdhouse.TextGrid", 
  package = "tjm.praat"
)
```

We also need a place to save the image. I'm going to use a temporary file.

```{r}
png_out <- tempfile("birdhouse", fileext = ".png")
```

Shoot. I just forgot what the arguments are to this script. That's okay, we 
can **print the function to view the script form**. 

```{r}
f_draw_textgrid

print(f_draw_textgrid, condense = FALSE)
```

Oh that's right. *Width* then *height*. Now, let's call the function and view
the resulting image.

```{r png-demo}
result <- f_draw_textgrid(tg_in, 7, 2, png_out)
magick::image_read(result) 
```

Note that when we printed the wrapped script's function, the first line of the
output (**the function signature**) included the names of the variables in the Praat
form. Indeed, the arguments to this function are set based on the Praat
form:

```{r}
args(f_draw_textgrid)
```

This feature means that a text's editor autocomplete/hint system can help us
remember the arguments to the Praat script and that we can reorder the arguments
in the function call as long as we use the correct names.

```{r}
result <- f_draw_textgrid(
  png_out = png_out, 
  height = 2,
  width = 7, 
  textgrid_in = tg_in
)
```

## Example using bundled Praat scripts

I have bundled some Praat scripts with this package. They are very minimal and
written under the assumption that the scripts would only ever called via a
Praat script. 

In one of my projects, I needed to extract the silences identified by a forced
alignment algorithm. Thus, I needed to do three things for each textgrid:

1.  Copy the tier with speech-sound intervals.
2.  Convert the speech-sound labels into "silence" and "sounding".
3.  Merged adjacent "silence" intervals and "sounding" intervals together.

So I wrote a script for each of these steps. The last-argument of each script is
a Praat textgrid, so I can pipe these wrapped-script functions into each other.

```{r}
f_duplicate <- wrap_praat_script(duplicate_tier)
f_relabel   <- wrap_praat_script(convert_tier_to_silences)
f_merge     <- wrap_praat_script(merge_duplicate_intervals)
```

Let's apply these scripts to our original example textgrid.

```{r png-demo-2}
tg_out <- tempfile("demo", fileext = ".TextGrid")
png_out <- tempfile("demo", fileext = ".png")

tg_result <- tg_in |> 
  f_duplicate("phones", "pauses", "last", tg_out) |> 
  f_relabel("pauses", "^$|sil|sp", tg_out) |>
  f_merge("pauses", tg_out)

png_result <- f_draw_textgrid(tg_result, 7, 2, png_out)

magick::image_read(png_result)
```

Now, with a little [purrr](https://purrr.tidyverse.org/) magic, we could run
this workflow on thousands of textgrids 😉.

Finally, as a little test, I want to make sure the package works when spaces
appear in the file names.

```{r png-demo-3}
tg_out <- tempfile("demo with spaces in name", fileext = ".TextGrid")
png_out <- tempfile("demo with spaces in name", fileext = ".png")

tg_result <- tg_in |> 
  f_duplicate("phones", "pauses", "last", tg_out) |> 
  f_relabel("pauses", "^$|sil|sp", tg_out) |> 
  f_merge("pauses", tg_out)

png_result <- f_draw_textgrid(tg_result, 3.5, 2, png_out)
basename(png_result)
```

## Acknowledgments

tjm.praat was created to process data from the [WISC Lab
project](https://kidspeech.wisc.edu/). Thus, development of this package was
supported by NIH R01DC009411 and NIH R01DC015653.
